---
title: "Hockey Season 1718"
author: "Alex Tin"
date: "5/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Desktop/shinyproject/hockey1718")
```


Libraries
```{r}
library(shiny)
library(shinydashboard)
library(ggplot2)
library(dplyr)
library(data.table)
library(plotly)
library(data.table)
library(DT)
library(tidyverse)
library(corrplot)

```

Load all the necessary files (4 total)
```{r}

#data for team stats in regular season
team2018reg <- read_csv(file='./data/NHLReg.csv',col_names = TRUE)

#data for player stats in playoffs (basic) - convert to team stats
team2018playoff <- read_csv(file='./data/NHLPlayoffs_Basic.csv',col_names=TRUE)

#data for player stats in playoffs (advanced) - convert to team stats
team2018playoffadv <- read_csv(file='./data/NHLPlayoffs_Advanced.csv',col_names=TRUE)

# Historical team data - from hockey-reference
#https://www.hockey-reference.com/play-index/tpbp_finder.cgi
teamhistreg <- read_csv(file='./data/Team2010_2019.csv',col_names=TRUE)
teamhistreg$`Eventual Champion` <- as.factor(teamhistreg$`Eventual Champion`)

str(teamhistreg)
```

Global

```{r}
# Teams for each division
Atlantic = c('BOS','TBL','TOR','FLA','MTL','BUF','OTT','DET')
Metro = c('WSH','PHI','PIT','CAR','CBJ','NYI','NYR','NJD')
Central = c('STL','COL','DAL','WPG','NSH','MIN','CHI')
Pacific = c('VEG','EDM','CGY','VAN','ARI','ANA','LAK','SJS')


fifty = 50

#Add divisions for all datasets
teamhistreg <- teamhistreg %>% mutate(., Division = ifelse(Tm %in% Atlantic,'Atlantic', ifelse(Tm %in% Metro, 'Metropolitan', ifelse(Tm %in% Central, 'Central','Pacific'))))



```

2010 - 2019 Team Stats (Tables)

```{r}
# Plotting champions of each year
champs_byseason <- teamhistreg[teamhistreg$`Eventual Champion` ==1,]


#By Season
histreg_byseason = teamhistreg %>% group_by(., Season)

#Average historical value
avgstat_byseason = teamhistreg %>% group_by(., Season) %>% summarise_all(., mean)

#Average of each stat (without 2012-13 season)
avgstat_total = teamhistreg %>% filter(., !Season=='2012-13') %>% summarise_all(., mean)


#By Team
histreg_byteam = teamhistreg %>% group_by(., Tm)




# Graph 3: Breakdown Per season, all teams
reg_byseason = teamhistreg %>% filter(., Season == '2010-11')



# Avg Team stat



```

2010 - 2019 Regular Season Team Stats (Graphs) (to be modified by reactive)


```{r}

#By Season

ggplot(histreg_byseason, aes(x=Season, y=`CF%`)) + geom_boxplot(aes(color=MadePlayoffs))






# Graph X: Breakdown Per season, all teams (violin plot)
teamhistreg %>% ggplot(., aes(x=Season, y=`HIT`)) + geom_violin(aes(fill=MadePlayoffs))

#Graph 2: Stat metric  vs. entire league per season (compared with historical average(?))
league_average <- avgstat_total %>% select(., `CF%`)

histreg_byteam %>% filter(., Season == '2017-18') %>% ggplot(., aes(x=Tm, y=`oiSH%`)) + geom_point(aes(color=MadePlayoffs, shape=Division)) + geom_hline(yintercept= as.integer(league_average)) + theme(axis.text.x = element_text(angle = 90)) + geom_text(aes(label=Tm),nudge_y = 0.5,size=2)


#Ex 2:
league_average2 <- avgstat_total %>% select(., `PDO`)

histreg_byteam %>% filter(., Season == '2017-18') %>% ggplot(., aes(x=Tm, y=`PDO`)) + geom_point(aes(color=MadePlayoffs, shape=Division)) + geom_hline(yintercept= as.integer(league_average2)) + theme(axis.text.x = element_text(angle = 90))

#Graph 3: Regular season stat metric for playoff teams (by season)

histreg_byteam %>% filter(., Season == '2010-11' & MadePlayoffs==TRUE) %>% ggplot(., aes(x=Tm, y=`oiSH%`)) + geom_point(aes(color=factor(`Eventual Champion`))) + geom_hline(yintercept= as.integer(league_average)) + theme(axis.text.x = element_text(angle = 90)) + scale_color_discrete(name="Results", breaks=c('1','2','0'), labels=c('Champion','Runner-up','Eliminated'))

#All playoff teams over the years
histreg_byteam %>% filter(., MadePlayoffs==TRUE) %>% ggplot(., aes(x=Tm, y=`CF%`)) + geom_point(aes(color=factor(`Eventual Champion`))) + geom_hline(yintercept= as.integer(league_average)) + theme(axis.text.x = element_text(angle = 90)) + scale_color_discrete(name="Results", breaks=c('1','2','0'), labels=c('Champion','Runner-up','Eliminated'))



#Graph 4: oiSH% by CF% graph (default), selectinput to put in 
histreg_byteam %>% filter(., Season == '2017-18') %>% ggplot(., aes(x=`dZS%`, y=`CF`)) + geom_point(aes(color=MadePlayoffs, shape=Division)) + theme(axis.text.x = element_text(angle = 90))

## Need to compare to playoffs!!

#Graph 5: The importance of faceoffs (apparently not really) - only change season variable
histreg_byteam %>% filter(., Season == '2017-18') %>% ggplot(., aes(x=`FO%`, y=`CF%`)) + geom_point(aes(color=MadePlayoffs, shape=Division)) + theme(axis.text.x = element_text(angle = 90)) + geom_vline(xintercept=fifty) + geom_hline(yintercept=fifty) + geom_text(aes(label=Tm),nudge_x = 0.4,size=2) 
# +
#     annotate('text', x=53, y = 55,label='High wins, high output') +
#     annotate('text', x=53, y = 44,label='High wins, low output') +
#     annotate('text', x=47, y = 55,label='Low wins, high output') +
#     annotate('text', x=47, y = 44,label='Low wins, low output')
#Takeaway - it doesn't seem to matter too much whether you win FaceOffs


    
    
#Graph 6: Importance of goalies
    
histreg_byteam %>% filter(., Season == '2017-18') %>% ggplot(., aes(x=`oiSV%`, y=`dZS%`)) + geom_point(aes(color=MadePlayoffs, shape=Division)) + theme(axis.text.x = element_text(angle = 90)) + geom_text(aes(label=Tm),nudge_y = 0.5,size=2) + geom_hline(yintercept = 50)



# Takeaway - increased shot percentage 

```



Regular Season Tab

Regular Season Player Table 
```{r}

# selecting the stats to compare
playertbl = playerreg %>% select(
    .,
    'DftYr',
    'Age',
    'Seasons',
    'H-Ref Name',
    'Team',
    'GP',
    'G',
    'A',
    'CPP',
    'CriG',
    'CruG',
    'CruA1',
    'CruA2',
    'MGL'
)


#Adding Divisions to player stats
playertbl <- playertbl %>% mutate(., Division = ifelse(Team %in% Atlantic,'Atlantic', ifelse(Team %in% Metro, 'Metropolitan', ifelse(Team %in% Central, 'Central','Pacific'))))

```


Regular Season Team Comparison table
```{r}
#Player statistic comparison

#adding Divisions to team 
teamregtbl <- teamreg %>% mutate(., Division = ifelse(Team %in% Atlantic,'Atlantic', ifelse(Team %in% Metro, 'Metropolitan', ifelse(Team %in% Central, 'Central','Pacific'))))



```


Playoff Tab

Playoff stats
```{r}
###Manipulate colnames of stats

#add identifier as playoff stats ('playoff_'), and convert to lowercase
colnames(teamplayoff) <- paste0('playoff_',tolower(colnames(teamplayoff)))

#add identifier as playoff stats ('playoff_'), and convert to lowercase
colnames(teamplayoffadv) <- paste0('playoff_',tolower(colnames(teamplayoffadv)))

#remove duplicate column except for identifier to merge with teamplay
teamplayoffadvfin <- subset(teamplayoffadv,select=-c(playoff_rk,playoff_x3, playoff_player,playoff_age,playoff_tm,playoff_pos,playoff_gp))

#merge teamplayoffs
playofftotal <- bind_cols(teamplayoff, teamplayoffadvfin)

#add divisions
playofftotal <- playofftotal %>% mutate(., division = ifelse(playoff_tm %in% Atlantic,'Atlantic', ifelse(playoff_tm %in% Metro, 'Metropolitan', ifelse(playoff_tm %in% Central, 'Central','Pacific'))))




# Team statistics
# teamstat <- regtbl %>% group_by(., Division, Team) %>% summarise(., agg_goals = sum(G), agg_clean = sum(CPP))
# 
# teamstat
# teamstat %>% ggplot(aes(x=agg_goals,y=agg_clean,label=Team)) + geom_point(aes(color=Division))
# 
# 
# meangoals = mean(regtbl$agggoals)
# 
# teamstat %>% mutate(., difftoavg= lapply(agggoals, function(x) x-meangoals))
# 


#Provided by <a href="https://www.sports-reference.com/sharing.html?utm_source=direct&utm_medium=Share&utm_campaign=ShareTool">Hockey-Reference.com</a>: <a href="https://www.hockey-reference.com/leagues/NHL_2018.html?sr&utm_source=direct&utm_medium=Share&utm_campaign=ShareTool#stats_adv">View Original Table</a><br>Generated 5/8/2020.
```


Glossary

Rk -- Rank
Corsi (EV)
CF -- Corsi For at Even Strength
Shots + Blocks + Misses
CA -- Corsi Against at Even Strength
Shots + Blocks + Misses
CF% -- Corsi For % at Even Strength
CF / (CF + CA)
Above 50% means the team was controlling the
puck more often than not with this player on the
ice in this situation.
Fenwick (EV)
FF -- Fenwick For at Even Strength
Shots + Misses
FA -- Fenwick Against at Even Strength
Shots + Misses
FF% -- Fenwick For % at Even Strength
FF / (FF + FA)
Above 50% means the team was controlling the
puck more often than not with this player on the
ice in this situation.
PDO (EV)
oiSH% -- Team On-Ice Shooting Percentage at Even Strength
Shooting % while this player/team was on the ice
oiSV% -- Team On-Ice Save Percentage at Even Strength
Save % while this player/team was on the ice
PDO -- PDO at Even Strength
Shooting % + Save %
Zone Starts (EV)
oZS% -- Offensive Zone Start % at Even Strength
Offensive Zone Faceoffs / (Offensive Zone Faceoffs + Defensive Zone Faceoffs),
that took place while on the ice
dZS% -- Defensive Zone Start % at Even Strength
Defensive Zone Faceoffs / (Offensive Zone Faceoffs + Defensive Zone Faceoffs),
that took place while on the ice
Faceoff Wins
FOW -- Faceoff Wins at Even Strength
Faceoff Losses
FOL -- Faceoff Losses at Even Strength
FO% -- Faceoff Win Percentage at Even Strength
HIT -- Hits at Even Strength
Blocks
BLK -- Blocks at Even Strength


For use in explaining the advanced stats:
https://www.sports-reference.com/blog/2016/10/what-the-heck-is-corsi-a-primer-on-advanced-hockey-statistics/










Correlation matrix

```{r}

avgseasonstat = subset(avgseasonstat, select = -c(Tm, `Eventual Champion`))

avgseasonstat


corrseasonstat <- subset(avgseasonstat, select = -c(Season, Rk))
corrseasonstat/82
corrseasontable = cor(corrseasonstat)

```


Corrplot

```{r}

corrplot(corrseasontable)


```








